Localisation and Sensor Privacy Using the Extended Information Filter and Secure Weighted Aggregation\thanksref{footnoteinfo}

### 2020-08-24 main.tex
Corrected how the restriction on weights was defined - weights must all be the same in queries of the same time step t Added remark which explains the restriction and the meaning behind weights chosen by the aggregator (how this affects the leakage function) Added m and public parameters which will include H to the encryption scheme definition Added the equations for our scheme

@@

Encryption of a plaintext message $m \in \mathbb{Z}_N$, producing ciphertext $c \in \mathbb{Z}^{*}_{N^2}$, is computed by
\begin{equation}
    c = g^m r^N \pmod{N^2}
\end{equation}
for a randomly chosen $r \in \mathbb{Z}_{N}$. $r^N$ can be considered the noise term which hides the value $g^m \pmod{N^2}$, which due to the scheme construction, is an easily computable discrete logarithm. The decryption of a ciphertext is computed by
\begin{equation}
    m = \frac{L(c^\lambda\pmod{n^2})}{L(g^\lambda\pmod{n^2})} \pmod{n}
\end{equation}
where $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{n}$.

In addition to encryption and decryption, the following homomorphic functions are provided by the Paillier scheme. $\forall m_1,m_2 \in \mathbb{Z}_n$,
\begin{align}
    \mathcal{D}(\mathcal{E}(m_1)\mathcal{E}(m_2)\!\pmod{n^2}) &= m_1+m_2\!\pmod{n} \\
    \mathcal{D}(\mathcal{E}(m_1)g^{m_2}\!\pmod{n^2}) &= m_1+m_2\!\pmod{n} \\
    \mathcal{D}(\mathcal{E}(m_1)^{m_2}\!\pmod{n^2}) &= m_1m_2\!\pmod{n}\,.
\end{align}

@@

Key generation is computed by choosing two equal length and sufficiently large primes $p$ and $q$, and computing $N=pq$. Additionally, hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}$ is defined, and the public key is set to $(N, H)$. $n$ private keys are generated by choosing $sk_i,\,1\leq i \leq n$ uniformly from $\mathbb{Z}_{N^2}$ and distributing them to all users, while the last key is set as
\begin{equation}
    sk_0 = -\sum^{n}_{i=1}sk_i \pmod{N^2}\,,
\end{equation}
and sent to the aggregator.

@@

When all encryptions $c^{(t)}_{i},\,1\leq i \leq n$ are sent to the aggregator, private summation and decryption are computed by the functions
\begin{equation}
    c^{(t)} = H(t)^{sk_0}\prod^{n}_{i=1}c^{(t)}_{i} \pmod{N^2}
\end{equation}
and
\begin{equation}
    \sum^{n}_{i=1}m^{(t)}_{i} = \frac{c^{(t)}-1}{N}\,. \label{eqn:agg_decryption}
\end{equation}
Correctness follows from $\sum^{n}_{i=0}sk_i = 0$, and thus
\begin{align*}
    &H(t)^{sk_0}\prod^{N}_{i=1}c_{i,t} \pmod{n^2} \\
    \equiv &H(t)^{sk_0}\prod^{N}_{i=1}(n+1)^{m_{i,t}} H(t)^{sk_i} \pmod{n^2} \\
    \equiv &H(t)^{\sum^N_{j=0}sk_j} \prod^{N}_{i=1}g^{m_{i,t}} \pmod{n^2} \\
    \equiv &(n+1)^{\sum^N_{i=1}m_{i,t}} \pmod{n^2}
\end{align*}
removing all noise terms.

@@

To perform the protocol from Figure \ref{fig:agg_steps} in a secure manner, we must first describe the security properties we aim to achieve. Broadly speaking, we want a cryptographic scheme which allows a time-series of homomorphically computed linear combinations of encrypted weights, to be summed by a private aggregation scheme. That is, we do not want sensors to learn the navigator weights, while we do not want the navigator to learn individual weighted linear combinations. This can be summarised by the two informal security notions:
\begin{description}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any additional knowledge about the navigator weights $\omega_j,\,1\leq j \leq m$ from receiving only their encryptions from the current and previous timesteps, and the ability to encrypt plaintexts of their choice.
    \item[Private Linear Combination Aggregation] No colluding subset \textit{excluding} the aggregator gains additional information about the remaining sensor values to be weighted $x^{(t)}_{i,j},\,0\leq j\leq m$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $y_i$ from the current and previous timesteps. Any colluding subset \textit{including} the aggregator learns only the sum of all linear combinations weighted by weights of their choice $\sum^n_{i=n}\sum^m_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j$. Note that this implies that $\sum^m_{j=1}x^{(t)}_{1,j}$ may be learned by the colluding subset given weights $(1,0,\dots,0)$, but that individual sensor values $x^{(t)}_{i,j}$ remain private.
\end{description}

From these definitions, it is clear that weights encrypted by an IND-CPA secure encryption scheme are sufficient for the first requirement, while a scheme satisfying AO is not sufficient for the second. To formalise the second requirement, we define a novel encryption type ``Linear-Combination Aggregator Oblivious Encryption'' and an accompanying security game, which capture the additional weights and modified leakage of AO.

@@

\begin{description}
    \item[$\mathsf{Setup}(\kappa)$] On input of security paramater $\kappa$, generate the aggregator's public and private keys $pk_0$ and $sk_0$, and the remaining user private keys $sk_i,\,1\leq i \leq n$.
    \item[$\mathsf{Enc}(pk_0, \omega)$] The aggregator and users can encrypt a weight $\omega$ with the aggregator public key $pk_0$, and obtain the encryption $\mathcal{E}_{pk_0}(\omega)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1),\dots,\mathcal{E}_{pk_0}(\omega_m), x_{i,1},\dots,x_{i,m})$] At time $t$, user $i$ computes and obtains the encrypted linear combination $y_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x_{i,j}\omega_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, y_1,\dots,y_n)$] At time $t$, the aggregator computes the aggregation of linear combinations $\sum^{n}_{i=1}\sum^{m}_{j=1} x_{i,j}\omega_j$ using its private key $sk_0$.
\end{description}

@@

\begin{description}
    \item[Setup] The challenger runs the $\mathsf{Setup}$ algorithm and gives $pk_0$ to the attacker
    \item[Queries] The attacker can now perform encryptions or submit queries that are answered by the challenger. The three types of actions are:
    \begin{enumerate}
        \item \textit{Encryption:} The attacker chooses a weight $\omega$ and computes an encryption of $\omega$ under the aggregator's public key $pk_0$, obtaining $\mathcal{E}_{pk_0}(\omega)$.
        \item \textit{Combine Queries:} The attacker chooses a tuple $(i,t,\mathcal{E}_{pk_0}(\omega_1),\dots,\mathcal{E}_{pk_0}(\omega_m),x_{i,1},\dots,x_{i,m})$ for a unique pair $(i,t)$ and is given back the encryption of the linear combination $\mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x_{i,j}\omega_j)$ encrypted under both the aggregator public key $pk_0$ and the secret key $sk_i$.
        \item \textit{Compromise queries:} The attacker chooses $i$ and receives the secret key $sk_i$. The aggregator's secret key may also be compromised (when choosing $i=0$).
    \end{enumerate} 
    \item[Challenge] Next, the attacker chooses a time $t^*$, and a subset of users $U \subseteq S$ where $S$ is the complete set of users for which no combine queries for time $t^*$ and no compromise queries are made for the duration of the game. The attacker then chooses two series of tuples
    \begin{equation*}
        \langle(i,t^*,\mathcal{E}_{pk_0}(\omega_1),\dots,\mathcal{E}_{pk_0}(\omega_m),x_{i,1},\dots,x_{i,m})\,|\,i \in U\rangle
    \end{equation*}
    and
    \begin{equation*}
        \langle(i,t^*,\mathcal{E}_{pk_0}(\omega_1'),\dots,\mathcal{E}_{pk_0}(\omega_m'),x_{i,1}',\dots,x_{i,m}')\,|\, i \in U\rangle\,,
    \end{equation*}
    and gives them to the challenger. In the case that $0 \in U$ (\textit{i.e.} the aggregator is compromised) and $U = S$, it is additionally required that
    \begin{equation*}
        \sum^{n}_{i=1}\sum^{m}_{j=1} x_{i,j}\omega_j = \sum^{n}_{i=1}\sum^{m}_{j=1} x_{i,j}'\omega_j'\,.
    \end{equation*}
    The challenger then chooses a random bit $b \in \{1,0\}$ and returns encryptions 
    \begin{equation*}
        \langle\mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x_{i,j}\omega_j)\,|\,i\in U\rangle
    \end{equation*}
    if $b=0$, or 
    \begin{equation*}
        \langle\mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x_{i,j}'\omega_j')\,|\,i\in U\rangle
    \end{equation*}
    otherwise.
    \item[More Queries] The attacker can now perform more encryptions and submit queries, so long as the queries do not break the requirements in the Challenge stage. That is, $U \subseteq S$.
    \item[Guess] At the end, the attacker outputs a bit $b'$ and wins the game only if $b' = b$. The advantage of attacker $\mathcal{A}$ is defined as
    \begin{equation*}
        \mathsf{Adv}^{LCAO}(\mathcal{A}) \coloneqq \left\lvert \mathsf{Pr}[b'=b] - \frac{1}{2}\right\rvert\,.
    \end{equation*} 
\end{description}

@@

>>>>

\theoremstyle{remark}
\newtheorem{remark}{Remark}

@@

Encryption of a plaintext message $m \in \mathbb{Z}_N$, producing ciphertext $c \in \mathbb{Z}^{*}_{N^2}$, is computed by
\begin{equation}
    c = g^m r^N \pmod{N^2}
\end{equation}
for a randomly chosen $r \in \mathbb{Z}_{N}$. $r^N$ can be considered the noise term which hides the value $g^m \pmod{N^2}$, which due to the scheme construction, is an easily computable discrete logarithm. The decryption of a ciphertext is computed by
\begin{equation}
    m = \frac{L(c^\lambda\pmod{N^2})}{L(g^\lambda\pmod{N^2})} \pmod{N}
\end{equation}
where $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{N}$.

In addition to encryption and decryption, the following homomorphic functions are provided by the Paillier scheme. $\forall m_1,m_2 \in \mathbb{Z}_N$,
\begin{align}
    \mathcal{D}(\mathcal{E}(m_1)\mathcal{E}(m_2)\!\pmod{N^2}) &= m_1+m_2\!\pmod{N} \\
    \mathcal{D}(\mathcal{E}(m_1)g^{m_2}\!\pmod{N^2}) &= m_1+m_2\!\pmod{N} \\
    \mathcal{D}(\mathcal{E}(m_1)^{m_2}\!\pmod{N^2}) &= m_1m_2\!\pmod{N}\,.
\end{align}

@@

Key generation is computed by choosing two equal length and sufficiently large primes $p$ and $q$, and computing $N=pq$. Additionally, hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}$ is defined, and the public key is set to $(N, H)$. $n$ private keys are generated by choosing $sk_i,\,1\leq i\leq n$ uniformly from $\mathbb{Z}_{N^2}$ and distributing them to all users, while the last key is set as
\begin{equation}
    sk_0 = -\sum^{n}_{i=1}sk_i \pmod{N^2}\,,
\end{equation}
and sent to the aggregator.

@@

When all encryptions $c^{(t)}_{i},\,1\leq i \leq n$ are sent to the aggregator, private summation and decryption are computed by the functions
\begin{equation}
    c^{(t)} = H(t)^{sk_0}\prod^{n}_{i=1}c^{(t)}_{i} \pmod{N^2}
\end{equation}
and
\begin{equation}
    \sum^{n}_{i=1}m^{(t)}_{i} = \frac{c^{(t)}-1}{N}\,. \label{eqn:agg_decryption}
\end{equation}
Correctness follows from $\sum^{n}_{i=0}sk_i = 0$, and thus
\begin{equation*}
    \begin{split}
        &H(t)^{sk_0}\prod^{n}_{i=1}c_{i,t} \pmod{N^2} \\
        \equiv &H(t)^{sk_0}\prod^{n}_{i=1}(N+1)^{m_{i,t}} H(t)^{sk_i} \pmod{N^2} \\
        \equiv &H(t)^{\sum^n_{j=0}sk_j} \prod^{n}_{i=1}g^{m_{i,t}} \pmod{N^2} \\
        \equiv &(N+1)^{\sum^n_{i=1}m_{i,t}} \pmod{N^2}
    \end{split}
\end{equation*}
removing all noise terms.

@@

To perform the protocol from Figure \ref{fig:agg_steps} in a secure manner, we must first describe the security properties we aim to achieve. Broadly speaking, we want a cryptographic scheme which allows a time-series of homomorphically computed linear combinations of encrypted weights, to be summed by a private aggregation scheme. That is, we do not want sensors to learn the navigator weights, while we do not want the navigator to learn individual weighted linear combinations. This can be summarised by the two informal security notions:
\begin{description}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any additional knowledge about the navigator weights $\omega_j,\,1\leq j \leq m$ from receiving only their encryptions from the current and previous timesteps, and the ability to encrypt plaintexts of their choice.
    \item[Private Linear Combination Aggregation] No colluding subset \textit{excluding} the aggregator gains additional information about the remaining sensor values to be weighted $x^{(t)}_{i,j},\,0\leq j\leq m$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $y_i$ from the current and previous timesteps. Any colluding subset \textit{including} the aggregator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^n_{i=n}\sum^m_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j$.
\end{description}

\begin{remark}
    The notion of a leakage function including parameters from the aggregator requires extra care to be taken when giving its definition. Since an attacker may compromise the aggregator, they have control over the choice of these parameters, and therefore over the leakage function. We note that in the leakage function above, $\sum^n_{i=n}\sum^m_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by the colluding subset, \textit{e.g.} $\sum^m_{j=1}x^{(t)}_{1,j}$ given weights $(1,0,\dots,0)$, but that individual sensor values $x^{(t)}_{i,j}$ remain private due to the requirement that all sensors receive the same weights.
\end{remark}

From the informal definitions above, it is clear that weights encrypted by an IND-CPA secure encryption scheme are sufficient for the first requirement, while a scheme satisfying AO is not sufficient for the second. To formalise the second requirement, we define a novel encryption type ``Linear-Combination Aggregator Oblivious Encryption'' and an accompanying security game, which capture the additional weights and modified leakage of AO.

@@

\begin{description}
    \item[$\mathsf{Setup}(\kappa)$] On input of security paramater $\kappa$, generate public parameters $\mathsf{pub}$, number of weights $m$, the aggregator's public and private keys $pk_0$ and $sk_0$, and the remaining user private keys $sk_i,\,1\leq i \leq n$.
    \item[$\mathsf{Enc}(pk_0, \omega)$] The aggregator and users can encrypt a weight $\omega$ with the aggregator public key $pk_0$, and obtain the encryption $\mathcal{E}_{pk_0}(\omega)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1),\dots,\mathcal{E}_{pk_0}(\omega_m), x_{i,1},\dots,x_{i,m})$] At time $t$, user $i$ computes and obtains the encrypted linear combination $y_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x_{i,j}\omega_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, y_1,\dots,y_n)$] At time $t$, the aggregator computes the aggregation of linear combinations $\sum^{n}_{i=1}\sum^{m}_{j=1} x_{i,j}\omega_j$ using its private key $sk_0$.
\end{description}

@@

\begin{description}
    \item[Setup] The challenger runs the $\mathsf{Setup}$ algorithm and gives $\mathsf{pub}$, $m$ and $pk_0$ to the attacker
    \item[Queries] The attacker can now perform encryptions or submit queries that are answered by the challenger. The three types of actions are:
    \begin{enumerate}
        \item \textit{Encryption:} The attacker chooses a weight $\omega$ and computes an encryption of $\omega$ under the aggregator's public key $pk_0$, obtaining $\mathcal{E}_{pk_0}(\omega)$.
        \item \textit{Combine Queries:} The attacker chooses a tuple $(i,t,\mathcal{E}_{pk_0}(\omega_1),\dots,\mathcal{E}_{pk_0}(\omega_m),x_{i,1},\dots,x_{i,m})$ such that for any two chosen combine query tuples $(i,t,\mathcal{E}_{pk_0}(\omega_1),\dots,\mathcal{E}_{pk_0}(\omega_m),x_{i,1},\dots,x_{i,m})$ and $(i',t',\mathcal{E}_{pk_0}(\omega_1'),\dots,\mathcal{E}_{pk_0}(\omega_m'),x_{i',1}',\dots,x_{i',m}')$ the following conditions hold:
        \begin{equation*}
            t = t' \implies \mathcal{E}_{pk_0}(\omega_j) = \mathcal{E}_{pk_0}(\omega_j'),\,1\leq j\leq m
        \end{equation*}
        and
        \begin{equation*}
            i = i' \wedge t = t' \implies x_{i,m} = x_{i',m}',\,1\leq j\leq m\,.
        \end{equation*}
        They are given back the encryption of the linear combination $\mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x_{i,j}\omega_j)$ encrypted under both the aggregator public key $pk_0$ and the secret key $sk_i$.
        \item \textit{Compromise queries:} The attacker chooses $i$ and receives the secret key $sk_i$. The aggregator's secret key may also be compromised (when choosing $i=0$).
    \end{enumerate} 
    \item[Challenge] Next, the attacker chooses a time $t^*$, and a subset of users $U \subseteq S$ where $S$ is the complete set of users for which no combine queries for time $t^*$ and no compromise queries are made for the duration of the game. The attacker then chooses two series of tuples
    \begin{equation*}
        \langle(i,t^*,\mathcal{E}_{pk_0}(\omega_1),\dots,\mathcal{E}_{pk_0}(\omega_m),x_{i,1},\dots,x_{i,m})\,|\,i \in U\rangle
    \end{equation*}
    and
    \begin{equation*}
        \langle(i,t^*,\mathcal{E}_{pk_0}(\omega_1'),\dots,\mathcal{E}_{pk_0}(\omega_m'),x_{i,1}',\dots,x_{i,m}')\,|\, i \in U\rangle\,,
    \end{equation*}
    and gives them to the challenger. In the case that $0 \in U$ (\textit{i.e.} the aggregator is compromised) and $U = S$, it is additionally required that
    \begin{equation*}
        \sum^{n}_{i=1}\sum^{m}_{j=1} x_{i,j}\omega_j = \sum^{n}_{i=1}\sum^{m}_{j=1} x_{i,j}'\omega_j'\,.
    \end{equation*}
    The challenger then chooses a random bit $b \in \{1,0\}$ and returns encryptions 
    \begin{equation*}
        \langle\mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x_{i,j}\omega_j)\,|\,i\in U\rangle
    \end{equation*}
    if $b=0$, or 
    \begin{equation*}
        \langle\mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}x_{i,j}'\omega_j')\,|\,i\in U\rangle
    \end{equation*}
    otherwise.
    \item[More Queries] The attacker can now perform more encryptions and submit queries, so long as the queries do not break the requirements in the Challenge stage. That is, $U \subseteq S$.
    \item[Guess] At the end, the attacker outputs a bit $b'$ and wins the game only if $b' = b$. The advantage of attacker $\mathcal{A}$ is defined as
    \begin{equation*}
        \mathsf{Adv}^{LCAO}(\mathcal{A}) \coloneqq \left\lvert \mathsf{Pr}[b'=b] - \frac{1}{2}\right\rvert\,.
    \end{equation*} 
\end{description}

@@

Enccryption: Choose random $r$ and compute
\begin{equation}
    \mathcal{E}_{pk_0}(\omega) = (N+1)^{\omega}r^N \pmod{N^2}\,,
\end{equation}
Combine encryption:
\begin{equation}
    \begin{split}
        y^{(t)}_i &= \mathcal{E}_{pk_0,sk_i}\left(\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j\right) \\
        &= H(t)^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(t)}_j)^{x^{(t)}_{i,j}} \pmod{N^2} \\
        &= H(t)^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(x^{(t)}_{i,j}\omega^{(t)}_j) \pmod{N^2} \\
        &= H(t)^{sk_i}\prod^{m}_{j=1}(N+1)^{x^{(t)}_{i,j}\omega^{(t)}_j}r^N \pmod{N^2} \\
        &= H(t)^{sk_i}(N+1)^{\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j}r^N \pmod{N^2}
    \end{split}
\end{equation}
for a random $r$ (not the same as above)

Aggregation and decryption in two steps. Aggregation:
\begin{equation}
    \begin{split}
        y^{(t)} &= \prod^n_{i=1}y^{(t)}_i \pmod{N^2} \\
        &= \prod^n_{i=1}H(t)^{sk_i}(N+1)^{\sum^{m}_{j=1}x_{i,j}\omega_j}r^N \pmod{N^2} \\
        &= H(t)^{\sum^n_{i=1}sk_i}\prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j}r^N \pmod{N^2} \\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j}r'^N \pmod{N^2}
    \end{split}
\end{equation}
for some $r' \in \mathbb{Z}_N$.

Decryption:
\begin{equation}
\sum^{n}_{i=1}\sum^{m}_{j=1}x^{(t)}_{i,j}\omega^{(t)}_j = \frac{L((y^{(t)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}
\end{equation}
